\thispagestyle{empty}
\vspace*{12\baselineskip}

{\large\textbf{Acknowledgments:} }
\par

This thesis would have not been possible without the support I received from many individuals.

First and foremost, I would like to thank Prof.~Dr.-Ing.~Georg Carle for giving me the opportunity to join the Chair of Network Architectures and Services and for supervising the dissertation. Similarly, I am most thankful to Dr.-Ing.~Henning Sanneck for getting me in contact with Prof.~Carle, and for setting up and maintaining the framework which financed my position at the chair.

I am forever grateful to my parents Krisztina and G\'abor. Without their unquestioning and unwavering support to pursue my goals, I could not be who I am and where I am today.

Throughout my research, I have received invaluable guidance -- about science, life, and everything in-between -- from my friend, mentor and colleague Stephen Mwanje.
My band of brothers at the chair, fellow Ph.~D. students Richard and Filip, have been the greatest allies a man can wish for.
My friends and colleagues at Nokia, chief among them Christian and Christoph, have provided a helping hand and a listening ear to ease my troubles too many times to count.
Thank you all for your friendship and support, you have made this journey truly worth undertaking!

I am grateful to Afsaneh Gharouni for providing companionship and ease of mind in the hardest of times.

I would also like to thank my students who helped me and contributed to my research, especially my friend and co-author Janik Schnellbach.

Finally, I would like to thank Prof.~Dr.~Rofl Stadler for being my second assessor, and Prof.~Debarghya Ghoshdastidar for chairing the examination committee.

\vspace{20mm}

\hfill M\"unchen, 26.05.2023

\vspace{5mm}
\hfill M\'arton Kaj\'o


\clearpage






